---
title: "Acer_damage"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

AMH - Load packages, set your working directory, and add data AMY
```{r}
library(tidyverse)
library(ggpubr)
library(RColorBrewer)
library(rcompanion)


setwd("C:/Users/hrusk/Dropbox (Smithsonian)/2021_SummerInterns/EB") #amy laptop# 

damage <- read.csv("A_rubrum_damage_04Aug2021.csv")


```

EB - Packages, Working Directory, Data 
```{r}
library(tidyverse)
library(ggpubr)
library(RColorBrewer)
library(rcompanion)


setwd("~/Dropbox (University of Michigan)/2021_SummerInterns/EB")
damage <- read.csv("A_rubrum_damage_04Aug2021.csv")
```


Graphing function
```{r}
#barGraphStats(data=, variable="", byFactorNames=c(""))
barGraphStats <- function(data, variable, byFactorNames) {
  count <- length(byFactorNames)
  N <- aggregate(data[[variable]], data[byFactorNames], FUN=length)
  names(N)[1:count] <- byFactorNames
  names(N) <- sub("^x$", "N", names(N))
  mean <- aggregate(data[[variable]], data[byFactorNames], FUN=mean)
  names(mean)[1:count] <- byFactorNames
  names(mean) <- sub("^x$", "mean", names(mean))
  sd <- aggregate(data[[variable]], data[byFactorNames], FUN=sd)
  names(sd)[1:count] <- byFactorNames
  names(sd) <- sub("^x$", "sd", names(sd))
  preSummaryStats <- merge(N, mean, by=byFactorNames)
  finalSummaryStats <- merge(preSummaryStats, sd, by=byFactorNames)
  finalSummaryStats$se <- finalSummaryStats$sd / sqrt(finalSummaryStats$N)
  return(finalSummaryStats)
}
```

Data manipulation

```{r}

#income level to character#
damage$Income <- as.character(damage$Income)

##replace "NA"s in Forest Income Level##
damage$Income <- damage$Income %>% replace_na("None")


##remove NAs##
damage2<- na.omit(damage)

#make percent damage numeric variable#
damage2$Percent_damage <- as.numeric(damage2$Percent_damage)

##make column for herbivores##
damage2 <- damage2 %>%
  mutate(herbivore = Grasshoppers_crickets_katydids + Caterpillar_like + Hoppers + Aphids + Thrips + Herb_beetles + Whiteflies_mealybugs_scale.insects + Mirid + Larve_pinchers)

##make column for predators##
damage2 <- damage2 %>%
  mutate(predator = Pred_beetles + Assassins + Spiders + Wasps)

##create dataframe without forest##
damage3 <- damage2 %>%
  filter(Area_type == "Urban"|Area_type == "Suburban ")

```


Explore percent damage graphs

```{r}
## Percent damage vs. Area ##
EBpalette1 <- c("#00B2A9", "#989C97", "#D86018", "#702082")
  
g1 <- ggplot(data = barGraphStats(data = damage2, variable = "Percent_damage", byFactorNames = c("Area_type")), aes(x=Area_type, y=mean, fill=Area_type)) +
    geom_bar(stat='identity', position=position_dodge(), width=0.8) +
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width = 0.2, position = position_dodge(0.9)) +
scale_fill_manual(values = EBpalette1) + theme(legend.position = "none") + xlab("Area Type") + ylab("Percent Damage")

g11 <- g1 + theme(panel.background = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
```

Another way to graph (with a colorblind palette)
```{r}
##Creating colourblind palette###
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

##saved graph as variable to edit theme below##
g2 <- ggplot(data = barGraphStats(data = damage2, variable = "Percent_damage", byFactorNames = c("Area_type")), aes(x=Area_type, y=mean, fill = Area_type)) +
    geom_bar(stat='identity', position=position_dodge()) +
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width = 0.2, position = position_dodge(0.9)) +
    xlab("Area type") + ylab("Percent Damage") +
    scale_fill_manual(values = cbPalette) 

##setting theme to remove background grid and add axis lines##    
g22 <- g2 + theme(panel.background = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

g1 + theme(panel.background = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size = 24),axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)), axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)), legend.position = "none")
           
```

```{r}
## Percent damage vs. Area and Income Level ##

damage2$Income <- factor(damage2$Income, levels=c("Low", "Medium", "High", "None"))
EBpalette <- c("#FFCB05", "#00274C", "#2F65A7", "#00B2A9")

g3 <- ggplot(data = barGraphStats(data = damage2, variable = "Percent_damage", byFactorNames = c("Area_type", "Income")), aes(x= Area_type, y=mean, fill=Income)) +
    geom_bar(stat='identity', position=position_dodge()) +
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width = 0.2, position = position_dodge(0.9)) +

scale_fill_manual(values = EBpalette) + xlab("Area Type") + ylab("Percent Damage")

g33 <- g3 + theme(panel.background = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + xlab("Area type") + ylab("Percent Damage") +
    scale_fill_manual(values = EBpalette)

g2 + theme(panel.background = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), text = element_text(size = 24), axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)), axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)))


```

Income graph with legend 
```{r}
damage2$Income <- factor(damage2$Income, levels=c("Low", "Medium", "High", "None"))

g3 <- ggplot(data = barGraphStats(data = damage2, variable = "Percent_damage", byFactorNames = c("Area_type", "Income")), aes(x= Area_type, y=mean, fill=Income)) +
    geom_bar(stat='identity', position=position_dodge()) +
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width = 0.2, position = position_dodge(0.9)) +
scale_fill_manual()

```


Preparing for Analysis

Test for Norm: Spoiler Alert it is not 
```{r}
shapiro.test(damage2$Percent_damage)
ggqqplot(damage2$Percent_damage)

damage2$Percent_damage_t <- transformTukey(damage2$Percent_damage)
tukey_lamda <- transformTukey(damage2$Percent_damage, returnLambda =TRUE)

shapiro.test(damage2$Percent_damage_t)

damage2$Percent_damage_log <- log(damage2$Percent_damage +1)
shapiro.test(damage2$Percent_damage_log)

##Helping Visualize the Ugly that is my data##
plot(damage2$Percent_damage_log)
hist(damage2$Percent_damage_log)

```

Load in new packages 
```{r}

library(lme4)
library(nlme)

```

Ana<3
```{r}
#####mod1 <- lme(Percent_damage ~ Area_type, data = damage2, random = 1|Area_type/Tree_num)


library(lattice)
library(effects)

##Messing around with models
mod0 <- lm(Percent_damage ~ Area_type, data = damage2)

mod1 <- lme(Percent_damage ~Area_type, random = ~1|Area_type/Tree_num, data = damage2)

mod2 <- lme(Percent_damage ~ Area_type, random = ~1|Tree_num, data = damage2)

mod3 <- lme(Percent_damage ~ Area_type*Income, random = ~1|Tree_num, data = damage3)

anova(mod2)
anova(mod3)

```



```{r}
display.brewer.all(colorblindFriendly = TRUE)
install.packages("wesanderson")
library(wesanderson)

```{r}
##BUG GRAPHS

##Herb Area Type
g4 <- ggplot(data = barGraphStats(data = damage2, variable = "herbivore", byFactorNames = c("Area_type")), aes(x=Area_type, y=mean, fill=Area_type)) +
   geom_bar(stat='identity', position=position_dodge()) +
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width = 0.2, position = position_dodge(0.9)) +
scale_fill_manual(values = EBpalette1) + xlab("Area Type") + ylab("Herbivore Abundance") + theme(legend.position = "none")

g4 + theme(panel.background = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

##Pred Area Type
g5 <- ggplot(data = barGraphStats(data = damage2, variable = "predator", byFactorNames = c("Area_type")), aes(x=Area_type, y=mean, fill=Area_type)) +
    geom_bar(stat='identity', position=position_dodge()) +
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width = 0.2, position = position_dodge(0.9)) +
scale_fill_manual(values = EBpalette1) + xlab("Area Type") + ylab("Predator Abundance") + theme(legend.position = "none")

g5 + theme(panel.background = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

```


```{r}
##LM Scatterplot bugs

herbivore_lm <- lm(Percent_damage ~ herbivore, data = damage2)
summary(herbivore_lm)

predictedherbivore <- predict(herbivore_lm, interval = "confidence", level = 0.95)

ggplot(damage2, aes(x = herbivore, y = Percent_damage, color = Area_type)) + 
  geom_point(shape = 1) +
  scale_colour_hue(l=50) +
  geom_smooth(method = lm, se = FALSE, fullrange = TRUE) +
  xlab("Number of Herbivores") + ylab("Percent Damage")


```
>>>>>>> 2f10b395d2ce73fa7fbef14d9bdff1ee7a0a3f85

        
        
```
```{r}

full_dataherbivore <- data.frame(damage2, predictedherbivore)

g8 <- ggplot(full_dataherbivore, aes(x = herbivore, y = Percent_damage))+
  geom_point()+
  geom_line(aes(x = herbivore, y = fit)) +
  geom_ribbon(aes(x = herbivore, ymin = lwr, ymax = upr), alpha = 0.3) +
  annotate("text", x = 14, y = 60, label = "y = -1.68x + 22.99")+
  annotate("text", x = 14, y = 55, label = "italic(R) ^ 2 == 0.28", parse = TRUE)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))+
  labs(y = expression(Mean~number~of~entry~holes~on~cutting), x = expression(Distance~to~surface~water~(m) ))


ggplot(damage2, aes(x=herbivore, y=Percent_damage)) +
  geom_point(aes(color=Area_type)) + 
  scale_color_manual(values = c("purple",  "blue", "yellow")) +
  geom_smooth(aes(color=Area_type, fill=Area_type), size=1.35, method=lm) +
  labs(x="Herbivore Abundance", y="Perccent Damage") +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))
      
```
